A Fast Unified Model for Parsing and Sentence Understanding

### 2016-03-18 writing/hard_stack_paper/paper.tex
JG meta-edits, plus figure placement.

We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce\ operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard recurrent neural network sequence model. While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} We also include speed results from a simple C++/CUDA implementation of a recurrent neural network.

@@

Figure~\ref{fig:speed} compares the sentence encoding speed of the three models on random input data. Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size. At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--25x speedup relative to the standard CPU implementation. The thin-stack implementation is consistently about 4x slower than the recurrent neural network we tested here.

>>>>

We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce\ operation at any given time. The thin-stack algorithm (Algorithm~\ref{alg:thin-stack}) describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.

@@

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard RNN sequence model. While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT and our RNN baseline in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the optimized C++/CUDA models and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}

@@

Figure~\ref{fig:speed} compares the sentence encoding speed of the three models on random input data. Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size. At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), our model 10--25x faster than the standard CPU implementation, and about 4x slower than the RNN baseline.
